Visual Odometry for Autonomous Underwater Vehicles

### 2013-03-19 paper.tex
Fixed: fovis data is now correct

% author names and affiliations
\author{\IEEEauthorblockN{Stephan Wirth}
\IEEEauthorblockA{System, Robotics and Vision Group\\
Universitat de les Illes Balears\\
Palma de Mallorca, Spain\\
Email: stephan.wirth@uib.es}
\and
\IEEEauthorblockN{Pep Lluis Negre Carrasco}
\IEEEauthorblockA{System, Robotics and Vision Group\\
Universitat de les Illes Balears\\
Palma de Mallorca, Spain\\
Email: pl.negre@uib.cat} 
\and
\IEEEauthorblockN{Gabriel Oliver Codina}
\IEEEauthorblockA{System, Robotics and Vision Group\\
Universitat de les Illes Balears\\
Palma de Mallorca, Spain\\
Email: goliver@uib.es}}

@@

Visual odometry is the process of estimating a vehicle 3D pose using only visual images. This technique is becoming popular in AUVs for navigation, station keeping and the provision of a manipulator feedback information. Visual odometry output is often fused with Inertial Measurement Units (IMU) to provide a cheaper alternative to DVLs \cite{Hildebrandt2010}. However, the limitations of underwater vision are widely known, and its performance depends on many factors such as visibility, lighting and distorsion resulting from varying refractive indices.

@@

TODO: Any publications on underwater visual odometry (including AUVs), else UAV or automobile systems

@@

TODO Error comparison (table)

>>>>

% author names and affiliations
% \author{\IEEEauthorblockN{Stephan Wirth}
% \IEEEauthorblockA{System, Robotics and Vision Group\\
% Universitat de les Illes Balears\\
% Palma de Mallorca, Spain\\
% Email: stephan.wirth@uib.es}
% \and
% \IEEEauthorblockN{Pep Lluis Negre Carrasco}
% \IEEEauthorblockA{System, Robotics and Vision Group\\
% Universitat de les Illes Balears\\
% Palma de Mallorca, Spain\\
% Email: pl.negre@uib.cat} 
% \and
% \IEEEauthorblockN{Gabriel Oliver Codina}
% \IEEEauthorblockA{System, Robotics and Vision Group\\
% Universitat de les Illes Balears\\
% Palma de Mallorca, Spain\\
% Email: goliver@uib.es}}

\author{\IEEEauthorblockN{
Stephan Wirth, 
Pep Lluis Negre Carrasco, 
Gabriel Oliver-Codina}
\IEEEauthorblockA{
Departament de Matemàtiques i Informàtica, 
Universitat de les Illes Balears\\
Cra. Valldemossa, km. 7,5 07122 Palma de Mallorca (Spain)\\
Email: \{stephan.wirth, pl.negre, goliver\}@uib.es}}

@@

Visual odometry is the process of estimating a vehicle 3D pose using only visual images. This technique is becoming popular in AUVs for navigation, station keeping and the provision of a manipulator feedback information. Visual odometry output is often fused with Inertial Measurement Units (IMU) to provide a cheaper alternative to DVLs \cite{Hildebrandt2010} \cite{Dunbabin2004}. However, the limitations of underwater vision are widely known, and its performance depends on many factors such as visibility, lighting and distorsion resulting from varying refractive indices.

@@

The basic visual odometry algorithm pipeline \cite{Moravec1980} consists in identifying keypoints (landmarks) in each camera frame, extract the feature descriptors of these points, estimate the depth for every landmark (typically using stereo or rgb-d cameras), match features across time frames, and then estimate the rigid body transformation that best aligns the features between frames. The result of this process is an estimation of  camera motion between frames, therefore it is necessary to integrate this data over the time to obtain the vehicle position and orientation.

Visual odometers for AUVs \cite{Botelho2009} tend to use this pipeline and adapt every stage options (such us keypoint detector type and feature matching method) to they needs. Common feature detectors used for real-time visual odometry include Harris corners \cite{Harris1988} and FAST features \cite{Rosten2006}, which have a low computation cost and are robust against small viewpoint changes.

Several AUVs localization algorithms are based on mosaics or structured enviroments and therefore assumes prior knowledge of the area to explore \cite{Garcia2001} \cite{Gracias2003} \cite{Carreras2003}.

TODO

@@

\begin{table}[!t]
  \renewcommand{\arraystretch}{1.3}
  \caption{Comparison of translational and rotational errors for the laboratory experiments.}
  \label{tab:results_attr}
  \centering
  \begin{tabular}{|l|l|l|l|} \hline
    \textbf{Algorith} & \textbf{Translation} & \textbf{Rotation} & \textbf{Runtime}\\ \hline
    libviso2 & 0.00978\% & 0.001739 [deg/m] & TODO \\ \hline
    fovis    & 0.00601\% & 0.006311 [deg/m] & TODO \\ \hline
  \end{tabular} 
\end{table}

\begin{table}[!t]
  \renewcommand{\arraystretch}{1.3}
  \caption{Comparison of translational and rotational errors for the harbour experiments.}
  \label{tab:results_attr}
  \centering
  \begin{tabular}{|l|l|l|l|} \hline
    \textbf{Algorith} & \textbf{Translation} & \textbf{Rotation} & \textbf{Runtime}\\ \hline
    libviso2 & 0.00833\% & 0.007913 [deg/m] & TODO \\ \hline
    fovis    & 0.01049\% & 0.003122 [deg/m] & TODO \\ \hline
  \end{tabular} 
\end{table}
